---
title: "PSTAT127 Homework 4"
author: "Kevin Ayala"
date: "2/22/2019"
output: pdf_document
---
1)

(e) Fit a Poisson response model for the number of incidents with the predictors: log of service, type, year and period. Test whether the parameter associated with the service term can be one. Explain why we are interested in such a test
```{r}
library(MASS)
?ships
ships
library(tidyverse)
cleanships <- ships %>% filter(service != 0)
modelfit1 <- glm(incidents ~ log(service) + type + year + period, data = cleanships, family = poisson)
summary(modelfit1)

```


When we inspect the data closley, we notice that the log of the service is in a close enough range to 1, it is possible to model our rate in which in this case is incidents. This is becuase were hold constant the count response by using the Poisson regression while keeping the coefficeint with offset. Thus incident damage is correlated to service by the data upon further inspection. 

(f) Fit the Poisson rate model with all two-way interactions of the three predictors. Does this model fit the data?
```{r}
modelfit2 <- glm(incidents ~ (type + year + period)^2, data = cleanships, family = poisson(link = "log"), offset = log(service))
modelfit2
summary(modelfit2)

```

Yes it does fit the model, no predictors need to be dropped as non are significant as the p value is very close to 1 or is 1, which means we always reject the null hypothesis

(h) Now fit the rate model with just the main effects and compare it to the interaction model. Which model is preferred?
```{r}
modelfit3 <- glm(incidents ~ period + year + type, family = poisson(link = "log"),
            data = cleanships, offset = log(service))
summary(modelfit3)


```

Here we can use the Akaike Information Criteria and compare the 2nd and 3rd model. Under the sencond model we have it so that our AIC is 165 whereas this new third model has a AIC scoore of 146, which is smaller than the second. Thus the third model is superior under the Akaike information criterian. 

(i) Fit quasi Poisson versions of the two previous models and repeat the comparison.

```{r}
#model with no interaction effects 
modelfit4 <- glm(incidents ~ period + year + type, family = quasipoisson(link = "log"),
            data = cleanships, offset = log(service))
summary(modelfit4)

#model with interacton effects
modelfit5 <-  glm(incidents ~ (type + year + period)^2, data = cleanships, family = quasipoisson(link = "log"))
summary(modelfit5)

#model comparison
anova(modelfit4, modelfit5,test = "F")

```
With observe a p-value of .2454 between our comparisons of our two quasi poisson models, this indicates that at a .05 alpha level, we fail to reject the null hyphothesis of the main effects models being better. Thus we concolude main effects "modelfit4" quasi poisson is prefered. We note that there exists

(j) Interpret the coefficients of the main effects of the quasi-Poisson model. What factors are associated with higher and lower rates of damage incidents?


```{r}
#for coefficients tpe b and e
exp(0.32558 - (-0.54334))
#periods
exp(0.38447)
```

Given the information above, we observe that boates that are of type B and have lower indident rates compared to of those that are from type E and D. Based on the data, we know that that type E boats are 2.38 (rounded value) or about twice as likely to get into an incident that ships of type B. 

we observe that the rate of incident increases by 1.467, meaning that ships built after 1964 and before 1974 have higher chance of incident, where ships built before have lower insident rates. This is perhaps because older ships were perhaps easier to navigate/maintian since the technology was well known, wherease newer ships with newer tech are harder to maintain since not many people have expertise with recent tech by nature. 


2)
```{r}
 # not the same as the S-PLUS dataset
select <- MASS::select  #needed to define select here since tidyverse and Mass interfere with each other in select
longley 
names(longley)[1] <- "y"
lm.ridge(y ~ ., longley)
plot(lm.ridge(y ~ ., longley,
              lambda = seq(0,0.1,0.001)))
select(lm.ridge(y ~ ., longley,
               lambda = seq(0,0.1,0.0001)))


               
```

(a) Write the model that is being fitted (with assumptions).

$$Yi$$ = the ith observation for GNP implicit price deflator (1954=100)
where $$i = 1, . . . , 16$$, for ith obs/row
 
 The Gross National Prouct, (GNP), is denoted by $$x_{i2}$$
 The nunmber of unemployed (unemployed) is denoted by $$x_{i3}$$
 Number of people in armed forces (Armed.Forces) is denoted by $$x_{i4}$$
 noninstitutionalized’ population greater or equal to 14 years of age. (population) denoted by $$x_{i5}$$
 the year (time) as (Year) denoted by $$x_{i6}$$
 The numberof people emoployed (Employed) denoted by $$x_{i7}$$
 
Our regression model tries to predict/model the number of people employed (Employed), thus the model take the form
$$Y_i = \beta_0 + \sum_{j=1}^7 B_jx_{ij} + \epsilon_i$$ where $\epsilon \sim N(0,\sigma^2)$ are iid which follows a normal distribution


(b) Write a brief explanation of the patterns you observe in this plot, as $$\lambda$$ changes, relative to the OLS estimators.

We notice that the estimated coeffeint converges to 0 as $$\lambda$$ approaches $$\infty$$. Why?

When we observe $$\hat{\beta}$$ (red dashed line and black line) that it gets closer and closer to zero as the value of $$\lambda$$ increases. This is because the estimates of Beta ridge gets smaller than OLS estimatros. Inversely, we notice that $$\hat{\beta_{ridge}}$$ in pink and blue get bigger than OLS estimatres. We also notice that for the green line that as the value of $$\lambda$$ increases, that $\hat{\beta_{ridge}}$ is approximately equal to OLS estimates. 



